---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tibbleprint

Print data frames as tibbles.

## Installation

Install with:

``` r
remotes::install_github("moodymudskipper/tibbleprint")
```

## When to use / not to use

*{tibbleprint}* overrides `base::print.data.frame()` and 
`base::format.data.frame()`.

```{r example}
library(tibbleprint)
```

It's usually bad practice to override base methods. It's unlikely to have 
terrible consequences in this case though, and its nice to be able to benefit
from *{tibble}*'s pretty output on regular data frames.

In any case better use it on own projects, or remove from final deliverable to avoid confusing collaborator or reader, or make your local *R* authority grumpy.

It's also probably a bad idea to place the library call in your RProfile.

## How to use

After the package is attached, data frames will print like tibbles, and printing
of data frames will be affected by *{tibble}* and *{pillar}*'s options.

The only two differences are that the header is different and row names are displayed
by default when they exist. Tibbles don't have row names, so this adjustment was
necessary.

```{r cars}
mtcars

# the data didn't change
names(mtcars)
```

We can hide row names:

```{r}
print(mtcars, row.names = FALSE)
```

We can use any argument from `tibble:::print.tbl`:

```{r}
print(mtcars, n = 2)
```

The original base behavior is triggered by any of these conditions : 

* The `base` argument is set to `TRUE`.
* The `base` argument is not set and option `"tibbleprint.base"` is set to `TRUE`.
* Other arguments from `base::print.data.frame` 
  ("digits", "quote", "right", or "max") are used.

```{r}
options(tibbleprint.base = TRUE)
mtcars
options(tibbleprint.base = FALSE)
```

We made `n` work with base behavior for convenience.

```{r}
# digits is an argument of base::data.frame (so triggers base printing)
# n isn't, yet this works
print(mtcars, digits = 0, n = 3)
```

Tibbles usually print nicer than data.frames, an occasional exception is when you
want to display small objects nested inside list columns, in those case using
the `base` argument is useful.

```{r}
test <- data.frame(id = 1:2)
test$list_col <- list(c(x=1, y=2), c(x=2, y = 4))

test

print(test, base = TRUE)
```

## Aknowledgements

*{tibbleprint}* borrows a few essential lines of code from 
Kirill Müller and Hadley Wickham's *{tibble}* package, and the added value
comes essentially from their work.
